---
title: "Differences between No-/Appeal years"
format: html
editor: visual
---

```{r}
load(here("data", "assignment.RData"))

load(here('data', 'delivery_channel_crs_analysis.RData'))
load(here('data', 'em_dat_events.RData'))
```

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  rstatix,
  ggpubr
  )
```

-   We focus on disasters which led to a \*call for international assistance\*\* (proxy for the most extreme events). We'll call such events "appeal disasters".
-   We restrict the data to pairs of actors both registered in the CRS data set at year t.
-   we remove countries that have not experienced any appeal disaster over the period 2002-2022.

```{r}
# data
data_appeal <- copy(assignment_universe)
data_appeal <- data_appeal[, never_treated := sum(appeal_n, na.rm = TRUE),
                                   by = recipient_name
][never_treated != 0]
```

```{r}
# donors' characteristics
donor_chars <- unique(data_delivery[!(donor_code==75 & bi_multi=="Private Sector"),
                                    .(donor_code, bi_multi)])
# recipients' characteristics
recipient_chars <- unique(data_delivery[, .(
  recipient_code, region_name, incomegroup_name)])
```

```{r}
# join
data_appeal <- merge(data_appeal, donor_chars, by="donor_code")
data_appeal <- merge(data_appeal, recipient_chars, by="recipient_code")
```

# Recipient level

```{r}
recipient_data <- copy(data_appeal)
```

## Intensive margin

### Disbursements

```{r}
imarg_disb_data <- copy(recipient_data)
imarg_disb_data <- imarg_disb_data[, .(recipient_code, year, disbursement, treatment = appeal_n)
][, .(
  disb = sum(disbursement, na.rm = T)
), by = .(recipient_code, year, treatment)
][, disb_lag1 := shift(disb, type = "lag"), by = recipient_code
][, first_diff := disb - disb_lag1
][, var := (disb - disb_lag1)/disb_lag1 * 100, by = recipient_code
][, treatment := as_factor(treatment)]
head(imarg_disb_data)
```

```{r}
## Variation rate
# stats
stat <- summarize(imarg_disb_data,
                  mean=mean(var, na.rm = T),
                  median=median(var, na.rm = T),
                  .by = treatment)

# histogram
imarg_disb_data |> 
    ggplot(aes(x=var, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
imarg_disb_data |> 
  ggbetweenstats(
  x = treatment,
  y = var,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "Variation rate (%)",
  title = "Comparison of changes in disbursements",
  subtitle = "Recipient country-year level"
)

# table
tbl_summary(
  data = imarg_disb_data,
  by = treatment,
  include = var,
  label = list(var = "Variation rate (%)")
) |> 
  add_p()
```

> Difference (not statistically significant)

```{r}
## First-difference
# stats
stat <- summarize(imarg_disb_data,
                  mean=mean(first_diff, na.rm = T),
                  median=median(first_diff, na.rm = T),
                  .by = treatment)

# histogram
imarg_disb_data |> 
    ggplot(aes(x=first_diff, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
imarg_disb_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in disbursements",
  subtitle = "Recipient country-year level"
)

# table
tbl_summary(
  data = imarg_disb_data,
  by = treatment,
  include = first_diff,
  label = list(first_diff = "First-difference")
) |> 
  add_p()
```

> Difference (not statistically significant)

### Commitments

```{r}
imarg_commit_data <- copy(recipient_data)
imarg_commit_data <- imarg_commit_data[, .(recipient_code, year, commitment, treatment = appeal_n)
][, .(
  commit = sum(commitment, na.rm = T)
), by = .(recipient_code, year, treatment)
][, commit_lag1 := shift(commit, type = "lag"), by = recipient_code
][, first_diff := commit - commit_lag1
][, var := (commit - commit_lag1)/commit_lag1 * 100, by = recipient_code
][, treatment := as_factor(treatment)]
head(imarg_commit_data)
```

```{r}
## Variation rate
# stats
stat <- summarize(imarg_commit_data,
                  mean=mean(var, na.rm = T),
                  median=median(var, na.rm = T),
                  .by = treatment)

# histogram
imarg_commit_data |> 
    ggplot(aes(x=var, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
imarg_commit_data |> 
  ggbetweenstats(
  x = treatment,
  y = var,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "Variation rate (%)",
  title = "Comparison of changes in commitments",
  subtitle = "Recipient country-year level"
)

# table
imarg_commit_data |> 
  tbl_summary(
  by = treatment,
  include = var,
  label = list(var = "Variation rate (%)")
) |> 
  add_p()
```

> Difference (not statistically significant)

```{r}
## First difference
# stats
stat <- summarize(imarg_commit_data,
                  mean=mean(first_diff, na.rm = T),
                  median=median(first_diff, na.rm = T),
                  .by = treatment)

# histogram
imarg_commit_data |> 
    ggplot(aes(x=first_diff, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
imarg_commit_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in commitments",
  subtitle = "Recipient country-year level")

# table
imarg_commit_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff,
  label = list(first_diff = "First-difference")
) |> 
  add_p()
```

> Difference (not statistically significant)

## Extensive margin

```{r}
xmarg_data <- copy(recipient_data)
xmarg_data <- xmarg_data[, .(recipient_code, year, 
                             donor_code, sector_code, disbursement, 
                             treatment = appeal_n,
                             bi_multi)
][, treatment := as_factor(treatment)
][, .(
  nb_donor = uniqueN(donor_code[disbursement > 0]),
  nb_multi = uniqueN(donor_code[disbursement > 0 & bi_multi=="Multilateral"]),
  nb_bilat = uniqueN(donor_code[disbursement > 0 & bi_multi=="Bilateral"]),
  nb_sector = uniqueN(sector_code[disbursement > 0])
), by = .(recipient_code, year, treatment)
][, `:=` (
  nb_donor_lag1 = shift(nb_donor, type = "lag"),
  nb_multi_lag1 = shift(nb_multi, type = "lag"),
  nb_bilat_lag1 = shift(nb_bilat, type = "lag"),
  nb_sector_lag1 = shift(nb_sector, type = "lag")
), by = recipient_code
][, `:=` (
  first_diff_donor = nb_donor - nb_donor_lag1,
  first_diff_multi = nb_multi - nb_multi_lag1,
  first_diff_bilat = nb_bilat - nb_bilat_lag1,
  first_diff_sector = nb_sector - nb_sector_lag1
)]

head(xmarg_data)
```

### Donors

#### All

```{r}
# stats
stat <- summarize(xmarg_data,
                  mean=mean(first_diff_donor, na.rm = T),
                  median=median(first_diff_donor, na.rm = T),
                  .by = treatment)

# histogram
xmarg_data |> 
    ggplot(aes(x=first_diff_donor, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
xmarg_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_donor,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in active donors",
  subtitle = "Recipient country-year level")

# table
xmarg_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_donor,
  label = list(first_diff_donor = "Donors")
) |> 
  add_p()
```

> No significant difference.

#### Multilaterals

```{r}
# stats
stat <- summarize(xmarg_data,
                  mean=mean(first_diff_multi, na.rm = T),
                  median=median(first_diff_multi, na.rm = T),
                  .by = treatment)

# histogram
xmarg_data |> 
    ggplot(aes(x=first_diff_multi, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
xmarg_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_multi,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in active multilateral donors",
  subtitle = "Recipient country-year level")

# table
xmarg_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_multi,
  label = list(first_diff_multi = "Multilateral donors")
) |> 
  add_p()
```

> No difference

#### Bilaterals

```{r}
# stats
stat <- summarize(xmarg_data,
                  mean=mean(first_diff_bilat, na.rm = T),
                  median=median(first_diff_bilat, na.rm = T),
                  .by = treatment)

# histogram
xmarg_data |> 
    ggplot(aes(x=first_diff_bilat, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
xmarg_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_bilat,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in active bilateral donors",
  subtitle = "Recipient country-year level")

# table
xmarg_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_bilat,
  label = list(first_diff_bilat = "Bilateral donors")
) |> 
  add_p()
```

> No significant difference

### Sectors

```{r}
# stats
stat <- summarize(xmarg_data,
                  mean=mean(first_diff_sector, na.rm = T),
                  median=median(first_diff_sector, na.rm = T),
                  .by = treatment)

# histogram
xmarg_data |> 
    ggplot(aes(x=first_diff_sector, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
xmarg_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_sector,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in active sectors",
  subtitle = "Recipient country-year level")

# table
xmarg_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_sector,
  label = list(first_diff_sector = "Sectors")
) |> 
  add_p()
```

> No significant difference.

## Channel of delivery

```{r}
channel_data <- copy(recipient_data)
channel_data <- channel_data[, .(recipient_code, year, treatment = appeal_n,
                                 disbursement, channel_nonstate, channel_nonstate_ngo, 
                                 channel_nonstate_multi)
][, treatment := as_factor(treatment)
][, .( # share of disbursements through non-state (total, multi, ngo) channels
  nonstate_pct=100*sum(disbursement[channel_nonstate==1], na.rm=T)/sum(disbursement, na.rm=T),
  multi_pct=100*sum(disbursement[channel_nonstate_multi==1], na.rm=T)/sum(disbursement, na.rm=T),
  ngo_pct=100*sum(disbursement[channel_nonstate_ngo==1], na.rm=T)/sum(disbursement, na.rm=T)
), by = .(recipient_code, year, treatment)
][, `:=` ( # one-period lag
  nonstate_lag1 = shift(nonstate_pct, type = "lag"),
  multi_lag1 = shift(multi_pct, type = "lag"),
  ngo_lag1 = shift(ngo_pct, type = "lag")
), by = recipient_code
][, `:=` ( # first-difference
  first_diff_nonstate = nonstate_pct - nonstate_lag1,
  first_diff_multi = multi_pct - multi_lag1,
  first_diff_ngo = ngo_pct - ngo_lag1
)]
head(channel_data)
```

### Non-State

```{r}
# stats
stat <- summarize(channel_data,
                  mean=mean(first_diff_nonstate, na.rm = T),
                  median=median(first_diff_nonstate, na.rm = T),
                  .by = treatment)

# histogram
channel_data |> 
    ggplot(aes(x=first_diff_nonstate, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
channel_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_nonstate,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (% pts)",
  title = "Comparison of changes in the share of disbursements going through non-State channels of delivery",
  subtitle = "Recipient country-year level")

# table
channel_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_nonstate,
  label = list(first_diff_nonstate = "Non-State channels")
) |> 
  add_p()
```

> No significant change

### Multilaterals

```{r}
# stats
stat <- summarize(channel_data,
                  mean=mean(first_diff_multi, na.rm = T),
                  median=median(first_diff_multi, na.rm = T),
                  .by = treatment)

# histogram
channel_data |> 
    ggplot(aes(x=first_diff_multi, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
channel_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_multi,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (% pts)",
  title = "Comparison of changes in the share of disbursements\ngoing through Multilateral channels of delivery",
  subtitle = "Recipient country-year level")

# table
channel_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_multi,
  label = list(first_diff_multi = "Multilateral channels")
) |> 
  add_p()
```

> No significant change

### NGOs

```{r}
# stats
stat <- summarize(channel_data,
                  mean=mean(first_diff_ngo, na.rm = T),
                  median=median(first_diff_ngo, na.rm = T),
                  .by = treatment)

# histogram
channel_data |> 
    ggplot(aes(x=first_diff_ngo, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
channel_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_ngo,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (% points)",
  title = "Comparison of changes in the share of disbursements\ngoing through NGO channels of delivery",
  subtitle = "Recipient country-year level")

# table
channel_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_ngo,
  label = list(first_diff_ngo = "NGO channels")
) |> 
  add_p()
```

> Weak difference (not statistically significant)

# Donor-Recipient level

**/!\\ PROBLEM: TOO MANY ZEROES**

```{r}
donor_recipient_data <- copy(data_appeal)
```

## Intensive margin

### Disbursements

```{r}
imarg_disb_data <- copy(donor_recipient_data)
imarg_disb_data <- imarg_disb_data[, .(donor_code, recipient_code, year, disbursement, treatment = appeal_n)
][, treatment := as_factor(treatment)
][, .(
  disb = sum(disbursement, na.rm = T)
), by = .(donor_code, recipient_code, year, treatment)
][, disb_lag1 := shift(disb, type = "lag"), by = recipient_code
][, first_diff := disb - disb_lag1]

head(imarg_disb_data)
```

### Commitments

```{r}

```

## Extensive margin (sector)

```{r}
xmarg_data <- copy(donor_recipient_data)
xmarg_data <- xmarg_data[, .(donor_code, recipient_code, year, 
                             sector_code, disbursement, 
                             treatment = appeal_n,
                             bi_multi)
][, treatment := as_factor(treatment)
][, .(
  nb_donor = uniqueN(donor_code[disbursement > 0]),
  nb_multi = uniqueN(donor_code[disbursement > 0 & bi_multi=="Multilateral"]),
  nb_bilat = uniqueN(donor_code[disbursement > 0 & bi_multi=="Bilateral"]),
  nb_sector = uniqueN(sector_code[disbursement > 0])
), by = .(donor_code, recipient_code, year, treatment)
][, `:=` (
  nb_donor_lag1 = shift(nb_donor, type = "lag"),
  nb_multi_lag1 = shift(nb_multi, type = "lag"),
  nb_bilat_lag1 = shift(nb_bilat, type = "lag"),
  nb_sector_lag1 = shift(nb_sector, type = "lag")
), by = .(donor_code, recipient_code)
][, `:=` (
  first_diff_donor = nb_donor - nb_donor_lag1,
  first_diff_multi = nb_multi - nb_multi_lag1,
  first_diff_bilat = nb_bilat - nb_bilat_lag1,
  first_diff_sector = nb_sector - nb_sector_lag1
)]

head(xmarg_data)
```
