---
title: "Data analysis"
format: html
editor: visual
---

```{r load_packages}
source(here::here('scripts', 'library.R'))
```

-   Unit-level observation: `recipient`-`donor`-`year`-`sector`.
-   Restrict observations to actors in a dyad both registered (active) in the CRS data set at year t.

```{r load_data}
# load(here('data', 'data_oecd_crs_clean.RData'))
# load(here('data', 'data_emdat_clean.RData'))
load(here("data", "assignment.RData"))
```

```{r}
# data
data <- copy(assignment_universe)

# data_appeal <- data_appeal[, never_treated := sum(appeal_n, na.rm = TRUE), by = recipient_name][never_treated != 0]
```

```{r}
# donors' characteristics
donor_chars <- unique(data_delivery[!(donor_code==75 & bi_multi=="Private Sector"),
                                    .(donor_code, bi_multi)])
# recipients' characteristics
recipient_chars <- unique(data_delivery[, .(
  recipient_code, region_name, incomegroup_name)])
```

```{r}
# join
data_appeal <- merge(data_appeal, donor_chars, by="donor_code")
data_appeal <- merge(data_appeal, recipient_chars, by="recipient_code")
```

# Single-treatment only

-   Treatment status

```{r single_treated_recipient_data}

# select vars of interest
data_binary_treatment <- data[, .(recipient_name, year, treat_all = switching_event_dummy, treat_appeal = switching_appeal_dummy)]

# keep unique obs
data_binary_treatment.unique <- unique(data_binary_treatment)

# convert event dummy from factor to numeric
data_binary_treatment.unique$treat_all <- as.numeric(levels(data_binary_treatment.unique$treat_all))[data_binary_treatment.unique$treat_all] # all types of events

data_binary_treatment.unique$treat_appeal <- as.numeric(levels(data_binary_treatment.unique$treat_appeal))[data_binary_treatment.unique$treat_appeal] # only "appeal" events

# plot treatment status
panelView::panelview(1 ~ treat_all, # all event types
          data = data_binary_treatment.unique, index = c("recipient_name","year"), 
          xlab = "Year", ylab = "Country",
          legend.labs = c("Control Recipients", 
                          "Treated Recipients", 
                          "Missing"),
          main = "Treatment Status: All events",
          by.timing = TRUE,
          collapse.history = "TRUE")
# "appeal" events - binary
panelView::panelview(1 ~ treat_appeal, # 'appeal' event type only
          data = data_binary_treatment.unique, index = c("recipient_name","year"), 
          xlab = "Year", ylab = "Number of Countries",
          legend.labs = c("Control Recipients", 
                          "Treated Recipients", 
                          "Missing"),
          main = "Treatment Status: Events w/ call for international assistance",
          by.timing = TRUE,
          collapse.history = "TRUE")

```

-   Remove recipient countries with more than one treatment period (i.e. only keep countries with no or one treatment period).

```{r}

# sum number of treated years by recipient
data_all.treated_year_N <- data_binary_treatment.unique[, .( # all
  treated_year_N = sum(treat_all, na.rm = T)
), by = recipient_name
][order(-treated_year_N)]

data_appeal.treated_year_N <- data_binary_treatment.unique[, .( # appeal only
  treated_year_N = sum(treat_appeal, na.rm = T)
), by = recipient_name
][order(-treated_year_N)]

# remove recipients w/ more than one treated year
data_all.treated_year_N[treated_year_N > 1, ] # all 
data_appeal.treated_year_N[treated_year_N > 1, ] # appeal only

# create object with names of the recipients w/ more than one treated year
data_all.multi_treated_recipients <- data_all.treated_year_N[treated_year_N > 1, recipient_name] # all
data_appeal.multi_treated_recipients <- data_appeal.treated_year_N[treated_year_N > 1, recipient_name] # appeal only

# remove recipients w/ more than one treated year
data.single_treated_all <- data[!recipient_name %in% data_all.multi_treated_recipients, ] # all
data.single_treated_appeal <- data[!recipient_name %in% data_appeal.multi_treated_recipients, ] # appeal only

# list single treated recipients
unique(data.single_treated_all$recipient_name) # all
unique(data.single_treated_appeal$recipient_name) # appeal only
```

-   create new `period` variable: year - year\[treatment==1\], by = recipient

```{r}

# select relevant vars
between_periods_comparison <- single_treated_recipient_data[, .(
  recipient_code, recipient_name,
  year,
  commitment, disbursement,
  treatment = switching_appeal_dummy
)
# sum(marize) finance flows by recipient and year
][, .(
  commit = sum(commitment, na.rm = T),
  disb = sum(disbursement, na.rm = T)
), by = .(recipient_code, recipient_name, year, treatment)
# create new var `period` 
][, period := year - year[treatment==1], by = .(recipient_code, recipient_name)
# order observations by recipient and year
][order(recipient_code, year), 
# create new var `lag1`
][, lag1 := shift(commit, type = "lag"), by = .(recipient_code, recipient_name)
# create new var `diff`
][, diff := commit - lag1
]
  
head(between_periods_comparison)

data <- between_periods_comparison[period %in% c(-2:2)]

ggplot(aes(x = period, y = diff)) +
  geom_boxplot()

ggbetweenstats(
  data = data, 
  x = period,
  y = diff,
  type = "np"
)


data_delivery[recipient_name=="Philippines", .(year, recipient_name)][order(year), ][]
```

-   new variables lag1
-   new variables diff year - lag1
-   plot

# Recipient level

```{r}
recipient_data <- copy(data_appeal)
```

## Intensive margin

### Disbursements

```{r}
imarg_disb_data <- copy(recipient_data)
imarg_disb_data <- imarg_disb_data[, .(recipient_code, year, disbursement, treatment = appeal_n)
][, .(
  disb = sum(disbursement, na.rm = T)
), by = .(recipient_code, year, treatment)
][, disb_lag1 := shift(disb, type = "lag"), by = recipient_code
][, first_diff := disb - disb_lag1
][, var := (disb - disb_lag1)/disb_lag1 * 100, by = recipient_code
][, treatment := as_factor(treatment)]
head(imarg_disb_data)
```

```{r}
## Variation rate
# stats
stat <- summarize(imarg_disb_data,
                  mean=mean(var, na.rm = T),
                  median=median(var, na.rm = T),
                  .by = treatment)

# histogram
imarg_disb_data |> 
    ggplot(aes(x=var, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
imarg_disb_data |> 
  ggbetweenstats(
  x = treatment,
  y = var,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "Variation rate (%)",
  title = "Comparison of changes in disbursements",
  subtitle = "Recipient country-year level"
)

# table
tbl_summary(
  data = imarg_disb_data,
  by = treatment,
  include = var,
  label = list(var = "Variation rate (%)")
) |> 
  add_p()
```

> Difference (not statistically significant)

```{r}
## First-difference
# stats
stat <- summarize(imarg_disb_data,
                  mean=mean(first_diff, na.rm = T),
                  median=median(first_diff, na.rm = T),
                  .by = treatment)

# histogram
imarg_disb_data |> 
    ggplot(aes(x=first_diff, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
imarg_disb_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in disbursements",
  subtitle = "Recipient country-year level"
)

# table
tbl_summary(
  data = imarg_disb_data,
  by = treatment,
  include = first_diff,
  label = list(first_diff = "First-difference")
) |> 
  add_p()
```

> Difference (not statistically significant)

### Commitments

```{r}
imarg_commit_data <- copy(recipient_data)
imarg_commit_data <- imarg_commit_data[, .(recipient_code, year, commitment, treatment = appeal_n)
][, .(
  commit = sum(commitment, na.rm = T)
), by = .(recipient_code, year, treatment)
][, commit_lag1 := shift(commit, type = "lag"), by = recipient_code
][, first_diff := commit - commit_lag1
][, var := (commit - commit_lag1)/commit_lag1 * 100, by = recipient_code
][, treatment := as_factor(treatment)]
head(imarg_commit_data)
```

```{r}
## Variation rate
# stats
stat <- summarize(imarg_commit_data,
                  mean=mean(var, na.rm = T),
                  median=median(var, na.rm = T),
                  .by = treatment)

# histogram
imarg_commit_data |> 
    ggplot(aes(x=var, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
imarg_commit_data |> 
  ggbetweenstats(
  x = treatment,
  y = var,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "Variation rate (%)",
  title = "Comparison of changes in commitments",
  subtitle = "Recipient country-year level"
)

# table
imarg_commit_data |> 
  tbl_summary(
  by = treatment,
  include = var,
  label = list(var = "Variation rate (%)")
) |> 
  add_p()
```

> Difference (not statistically significant)

```{r}
## First difference
# stats
stat <- summarize(imarg_commit_data,
                  mean=mean(first_diff, na.rm = T),
                  median=median(first_diff, na.rm = T),
                  .by = treatment)

# histogram
imarg_commit_data |> 
    ggplot(aes(x=first_diff, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
imarg_commit_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in commitments",
  subtitle = "Recipient country-year level")

# table
imarg_commit_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff,
  label = list(first_diff = "First-difference")
) |> 
  add_p()
```

> Difference (not statistically significant)

## Extensive margin

```{r}
xmarg_data <- copy(recipient_data)
xmarg_data <- xmarg_data[, .(recipient_code, year, 
                             donor_code, sector_code, disbursement, 
                             treatment = appeal_n,
                             bi_multi)
][, treatment := as_factor(treatment)
][, .(
  nb_donor = uniqueN(donor_code[disbursement > 0]),
  nb_multi = uniqueN(donor_code[disbursement > 0 & bi_multi=="Multilateral"]),
  nb_bilat = uniqueN(donor_code[disbursement > 0 & bi_multi=="Bilateral"]),
  nb_sector = uniqueN(sector_code[disbursement > 0])
), by = .(recipient_code, year, treatment)
][, `:=` (
  nb_donor_lag1 = shift(nb_donor, type = "lag"),
  nb_multi_lag1 = shift(nb_multi, type = "lag"),
  nb_bilat_lag1 = shift(nb_bilat, type = "lag"),
  nb_sector_lag1 = shift(nb_sector, type = "lag")
), by = recipient_code
][, `:=` (
  first_diff_donor = nb_donor - nb_donor_lag1,
  first_diff_multi = nb_multi - nb_multi_lag1,
  first_diff_bilat = nb_bilat - nb_bilat_lag1,
  first_diff_sector = nb_sector - nb_sector_lag1
)]

head(xmarg_data)
```

### Donors

#### All

```{r}
# stats
stat <- summarize(xmarg_data,
                  mean=mean(first_diff_donor, na.rm = T),
                  median=median(first_diff_donor, na.rm = T),
                  .by = treatment)

# histogram
xmarg_data |> 
    ggplot(aes(x=first_diff_donor, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
xmarg_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_donor,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in active donors",
  subtitle = "Recipient country-year level")

# table
xmarg_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_donor,
  label = list(first_diff_donor = "Donors")
) |> 
  add_p()
```

> No significant difference.

#### Multilaterals

```{r}
# stats
stat <- summarize(xmarg_data,
                  mean=mean(first_diff_multi, na.rm = T),
                  median=median(first_diff_multi, na.rm = T),
                  .by = treatment)

# histogram
xmarg_data |> 
    ggplot(aes(x=first_diff_multi, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
xmarg_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_multi,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in active multilateral donors",
  subtitle = "Recipient country-year level")

# table
xmarg_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_multi,
  label = list(first_diff_multi = "Multilateral donors")
) |> 
  add_p()
```

> No difference

#### Bilaterals

```{r}
# stats
stat <- summarize(xmarg_data,
                  mean=mean(first_diff_bilat, na.rm = T),
                  median=median(first_diff_bilat, na.rm = T),
                  .by = treatment)

# histogram
xmarg_data |> 
    ggplot(aes(x=first_diff_bilat, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
xmarg_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_bilat,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in active bilateral donors",
  subtitle = "Recipient country-year level")

# table
xmarg_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_bilat,
  label = list(first_diff_bilat = "Bilateral donors")
) |> 
  add_p()
```

> No significant difference

### Sectors

```{r}
# stats
stat <- summarize(xmarg_data,
                  mean=mean(first_diff_sector, na.rm = T),
                  median=median(first_diff_sector, na.rm = T),
                  .by = treatment)

# histogram
xmarg_data |> 
    ggplot(aes(x=first_diff_sector, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
xmarg_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_sector,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (t - t-1)",
  title = "Comparison of changes in active sectors",
  subtitle = "Recipient country-year level")

# table
xmarg_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_sector,
  label = list(first_diff_sector = "Sectors")
) |> 
  add_p()
```

> No significant difference.

## Channel of delivery

```{r}
channel_data <- copy(recipient_data)
channel_data <- channel_data[, .(recipient_code, year, treatment = appeal_n,
                                 disbursement, channel_nonstate, channel_nonstate_ngo, 
                                 channel_nonstate_multi)
][, treatment := as_factor(treatment)
][, .( # share of disbursements through non-state (total, multi, ngo) channels
  nonstate_pct=100*sum(disbursement[channel_nonstate==1], na.rm=T)/sum(disbursement, na.rm=T),
  multi_pct=100*sum(disbursement[channel_nonstate_multi==1], na.rm=T)/sum(disbursement, na.rm=T),
  ngo_pct=100*sum(disbursement[channel_nonstate_ngo==1], na.rm=T)/sum(disbursement, na.rm=T)
), by = .(recipient_code, year, treatment)
][, `:=` ( # one-period lag
  nonstate_lag1 = shift(nonstate_pct, type = "lag"),
  multi_lag1 = shift(multi_pct, type = "lag"),
  ngo_lag1 = shift(ngo_pct, type = "lag")
), by = recipient_code
][, `:=` ( # first-difference
  first_diff_nonstate = nonstate_pct - nonstate_lag1,
  first_diff_multi = multi_pct - multi_lag1,
  first_diff_ngo = ngo_pct - ngo_lag1
)]
head(channel_data)
```

### Non-State

```{r}
# stats
stat <- summarize(channel_data,
                  mean=mean(first_diff_nonstate, na.rm = T),
                  median=median(first_diff_nonstate, na.rm = T),
                  .by = treatment)

# histogram
channel_data |> 
    ggplot(aes(x=first_diff_nonstate, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
channel_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_nonstate,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (% pts)",
  title = "Comparison of changes in the share of disbursements going through non-State channels of delivery",
  subtitle = "Recipient country-year level")

# table
channel_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_nonstate,
  label = list(first_diff_nonstate = "Non-State channels")
) |> 
  add_p()
```

> No significant change

### Multilaterals

```{r}
# stats
stat <- summarize(channel_data,
                  mean=mean(first_diff_multi, na.rm = T),
                  median=median(first_diff_multi, na.rm = T),
                  .by = treatment)

# histogram
channel_data |> 
    ggplot(aes(x=first_diff_multi, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
channel_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_multi,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (% pts)",
  title = "Comparison of changes in the share of disbursements\ngoing through Multilateral channels of delivery",
  subtitle = "Recipient country-year level")

# table
channel_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_multi,
  label = list(first_diff_multi = "Multilateral channels")
) |> 
  add_p()
```

> No significant change

### NGOs

```{r}
# stats
stat <- summarize(channel_data,
                  mean=mean(first_diff_ngo, na.rm = T),
                  median=median(first_diff_ngo, na.rm = T),
                  .by = treatment)

# histogram
channel_data |> 
    ggplot(aes(x=first_diff_ngo, fill=treatment)) +
    geom_histogram(alpha=.3) +
    geom_vline(data=stat, aes(xintercept = mean, color=treatment)) +
    geom_vline(data=stat, aes(xintercept = median, color=treatment), linetype="dotted") +
  theme_minimal()

# boxplot
channel_data |> 
  ggbetweenstats(
  x = treatment,
  y = first_diff_ngo,
  type = "np",
  results.subtitle = F,
  xlab = "# of 'appeal' disaster per year",
  ylab = "First difference (% points)",
  title = "Comparison of changes in the share of disbursements\ngoing through NGO channels of delivery",
  subtitle = "Recipient country-year level")

# table
channel_data |> 
  tbl_summary(
  by = treatment,
  include = first_diff_ngo,
  label = list(first_diff_ngo = "NGO channels")
) |> 
  add_p()
```

> Weak difference (not statistically significant)

# Donor-Recipient level

**/!\\ PROBLEM: TOO MANY ZEROES**

```{r}
donor_recipient_data <- copy(data_appeal)
```

## Intensive margin

### Disbursements

```{r}
imarg_disb_data <- copy(donor_recipient_data)
imarg_disb_data <- imarg_disb_data[, .(donor_code, recipient_code, year, disbursement, treatment = appeal_n)
][, treatment := as_factor(treatment)
][, .(
  disb = sum(disbursement, na.rm = T)
), by = .(donor_code, recipient_code, year, treatment)
][, disb_lag1 := shift(disb, type = "lag"), by = recipient_code
][, first_diff := disb - disb_lag1]

head(imarg_disb_data)
```

### Commitments

```{r}

```

## Extensive margin (sector)

```{r}
xmarg_data <- copy(donor_recipient_data)
xmarg_data <- xmarg_data[, .(donor_code, recipient_code, year, 
                             sector_code, disbursement, 
                             treatment = appeal_n,
                             bi_multi)
][, treatment := as_factor(treatment)
][, .(
  nb_donor = uniqueN(donor_code[disbursement > 0]),
  nb_multi = uniqueN(donor_code[disbursement > 0 & bi_multi=="Multilateral"]),
  nb_bilat = uniqueN(donor_code[disbursement > 0 & bi_multi=="Bilateral"]),
  nb_sector = uniqueN(sector_code[disbursement > 0])
), by = .(donor_code, recipient_code, year, treatment)
][, `:=` (
  nb_donor_lag1 = shift(nb_donor, type = "lag"),
  nb_multi_lag1 = shift(nb_multi, type = "lag"),
  nb_bilat_lag1 = shift(nb_bilat, type = "lag"),
  nb_sector_lag1 = shift(nb_sector, type = "lag")
), by = .(donor_code, recipient_code)
][, `:=` (
  first_diff_donor = nb_donor - nb_donor_lag1,
  first_diff_multi = nb_multi - nb_multi_lag1,
  first_diff_bilat = nb_bilat - nb_bilat_lag1,
  first_diff_sector = nb_sector - nb_sector_lag1
)]

head(xmarg_data)
```

------------------------------------------------------------------------
